Spatial phylogenetics of Japanese ferns: Patterns, processes, and implications for conservation

### 2021-09-21 ms/manuscript.Rmd
Finish fixing caption letters per AJB
Upper-case
No parentheses in figure

We used data compiled by @Ebihara2019b, which classifies each fern taxon as sexual, apomictic, mixed (both sexual and apomictic modes known from the same taxon), or unknown. 
We calculated the percentage of apomictic species in each grid-cell as the sum of apomictic and mixed taxa divided by total number of taxa (`r s_figure("envir")`b).

@@

The cleaned list of specimens included `r n_specimens` specimens representing `r n_specimens_taxa` taxa (`r n_specimens_taxa_percent`). 
The distribution of collection locations was highly skewed, with > 3,000 specimens per grid-cell northwest of Tokyo in Saitama prefecture, and many fewer elsewhere (`r s_figure("abundance")`a). 
Despite this, sampling redundancy was generally good [> 0.3; @Gonzalez-Orozco2014b] throughout the country (mean `r mean_redundancy`, final community data matrix; errors are SD unless mentioned otherwise) (`r s_figure("abundance")`b). 
The final community data matrix included `r ncol(comm_ferns)` taxa and `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` grid-cells. `r n_excluded_cells` grid-cells were excluded due to low redundancy (< 0.1); these were mostly from Hokkaido and coastal areas.

@@

Taxonomic richness is lowest on the northernmost island of Hokkaido, then increases with decreasing latitude until reaching a maximum in southern Kyushu, then decreases again in smaller islands further south (`r figure("raw-diversity")`a, `r s_figure("raw-lat")`a). 
Observed values of PD and FD were highly correlated with richness as expected (`r figure("raw-diversity")`b, c), showing an initial steep slope that gradually starts to level off at higher richness (`r s_figure("raw-biplot")`a, c). 
Observed values of RPD and RFD were not linearly related to richness; however, they showed high variance at low species richness, and low variance at high species richness (`r s_figure("raw-biplot")`b, d).

@@

Grid-cells with significantly low PD were almost all located on the main islands of Japan (Honshu, Kyushu, and Shikoku; `r figure("rand-diversity")`a). 
A smaller number of cells had significantly high PD, and these were located mostly on small southern (Ryukyu, Izu, Ogasawara) islands or coastal areas (Kii Peninsula). 
Grid-cells with significantly low RPD were mostly observed in the southern main islands, particularly in and around Kyushu (`r figure("rand-diversity")`b). 
Grid-cells with significantly high RPD are located both in the small southern islands and northern Honshu and Hokkaido. 
Results of the randomization test for functional diversity were broadly similar to phylogenetic diversity (`r figure("rand-diversity")`c, d).

@@

All models were significant relative to null models without any fixed effects (`r s_table("lrt-null")`). 
The effect of temperature dominated (had the greatest absolute *t*-value) in environmental models (`r figure("coeff")`a--e), and % apomictic taxa dominated in reproductive models (`r figure("coeff")`f--g); temperature and % apomictic taxa were significant in each model where they were included, whereas other predictors (area, precipitation, precipitation seasonality) were not significant in some models. 
Richness showed a hump-shaped relationship with temperature, reaching maximum richness at intermediate temperatures (`r figure("modelfit")`a). 
Phylogenetic and functional diversity showed either moderately (SES of PD, SES of FD) to strongly (SES of RPD, SES of RFD) hump-shaped relationships with temperature (`r figure("modelfit")`b--e). 
SES of RPD showed a clear, linear decline with % apomictic taxa (`r figure("modelfit")`g); a similar pattern was detected in SES of PD, but with wider confidence intervals (`r figure("modelfit")`f). 
Conditional AIC was lower (indicating better model fit) in the reproductive model compared to the environmental model for SES of RPD and SES of PD (`r s_table("aic")`).

@@

Clustering analysis grouped grid-cells into `r english2(k_taxonomy$optimal$k)` or `r english2(k_phylogeny$optimal$k)` bioregions (`r s_figure("k-plot")`) on the basis of taxonomic or phylogenetic distances, respectively. 
The vast majority of grid-cells fell to a subset of these bioregions: using taxonomic distances, they include bioregion 1 (Hokkaido and northern Honshu), bioregion 2 (southern Honshu, Kyushu, and Shikoku), bioregion 3 (Ryukyu Islands), and bioregion 4 (Ogasawara Islands) (`r figure("bioregions")`a; `r s_figure("dendrogram")`a). 
Phylogenetic distances produced similar results, but the Ryukyu and Ogasawara Islands merged into a single bioregion (3), and bioregion 1 extended slightly further south (`r figure("bioregions")`b; `r s_figure("dendrogram")`b). 
The other much smaller bioregions ("singletons" including only a single grid-cell each) are likely artifacts of insufficiently low taxon richness for robust clustering, and are not discussed further.

@@

The percentage of protected area in grid-cells with significantly high biodiversity range from `r min_protected_total` to `r max_protected_total` total, including `r min_protected_high` to `r max_protected_high` with high status and `r min_protected_med` to `r max_protected_med` with medium status (`r figure("conserv-status")`a). 
Cells with significantly high biodiversity had a greater percentage of area protected than the baseline rate for Japan at the medium protection status (`r scales::percent(total_percent_protected_medium, accuracy = 0.1)`) across all biodiversity metrics.
In contrast, cells with significantly high biodiversity had a lower percentage of area protected than the baseline rate for Japan at the high protection status (`r scales::percent(total_percent_protected_high, accuracy = 0.1)`) for most biodiversity metrics.
Only cells with significantly high PE had a higher percentage area protected (`r best_protected_high` high, `r best_protected_med` medium, `r best_protected_total` total) than the baseline rate at both levels of protection.

The range of Japanese deer in Japan increased from `r deer_overall[["1978"]]$area` (`r deer_overall[["1978"]]$percent`) in 1978 to `r deer_overall[["2003"]]$area` (`r deer_overall[["2003"]]$percent`) in 2003, and may be as high as `r deer_overall$estimated$area` (`r deer_overall$estimated$percent`) according to model estimates.
The percentage of area occupied by deer in grid-cells with significantly high biodiversity increased from `r deer_in_high_biodiv[["1978"]]$percent` in 1978 to `r deer_in_high_biodiv[["2003"]]$percent` in 2003 when averaged across biodiversity types.
Cells with high FD, PD, or PE tended to have less deer occupation compared to the baseline rate (`r figure("conserv-status")`b).
However, cells with high species richness consistently exceeded the baseline rate of deer presence, with up to `r deer_highest_danger$estimated$percent` occupied by deer in the estimated range dataset.

@@

The latitudinal hump-shaped pattern of species richness observed in ferns here (Figs. `r figure_num("raw-diversity")`a, `r s_figure_num("raw-lat")`a) is well-documented [@Kusumoto2017; @Ebihara2019b], and similar patterns have been observed in deciduous broadleaved trees and conifers [@Shiono2015], and niche-modeled distributions of woody plants in Japan [@Fukaya2020]. 
Our spatial modeling analysis indicates that this is strongly linked to temperature, with maximum richness occurring at intermediate temperatures (`r figure("modelfit")`a).
Previous studies have identified temperature, water availability [@Bickford2006; @Qian2012], or a combination of these [@Nagalingum2015; @Link-Perez2018; @Qian2021] to be important drivers of spatial fern diversity.
A hump-shaped pattern of richness along elevation (maximum richness at intermediate elevations) has been frequently observed in ferns on mountains in tropical or sub-tropical areas [@Kessler2011a; @Suissa2021]. 
Hypotheses to explain such a pattern include purely geometrical constraints as well as biological hypothesis positing maximally favorable environmental conditions at mid-elevations [@Colwell2004a]. 
Another reason for the greater richness on the main islands compared to southern islands in Japan is that the main islands harbor higher elevational variation that may support a wider range of taxa than is present in the southern islands.

The overwhelming pattern characterizing phylogenetic and functional diversity (PD and FD) is the high diversity of southern islands (Ryukyu and Ogasawara Isl.) (`r figure("rand-diversity")`a, c). 
This is almost certainly due to the presence of primarily tropical lineages that do not occur in other parts of the country; `r n_southern_genera` genera and `r n_southern_families` families only occur south of 30.1° latitude (just south of Yakushima Isl.). 
The similarity in patterns of PD and FD is likely due to the moderate degree of phylogenetic signal present in the traits used in our analysis (Tables `r s_table_num("phy-sig")`, `r s_table_num("phy-sig-binary")`), and suggests that in ferns, phylogeny can be used as a reasonable stand-in for functional diversity. 
Given their high phylogenetic diversity, it is perhaps not surprising that southern islands also host high amounts of phylogenetic endemism; indeed, the vast majority of cells with significant levels of PE, as well as paleoendemic cells, are located in the southern islands (Figs. `r figure_num("endemism")`, `r s_figure_num("endemism-restricted")`). 
This is clearly due to the small land area of these islands, which are the only areas with a subtropical climate in Japan.

Another striking pattern revealed here is the predominance of long branches in the northern half of Honshu and the southern islands, and the predominance of short branches in southern Honshu, Shikoku, and Kyushu (`r figure("rand-diversity")`b). 
The long branches in the southern islands are likely being driven by the tropical lineages mentioned previously; those in the north may be due to refugial lineages that are distantly related to others and become more species rich at high-latitudes, such as *Equisetum* (`r s_figure("genus-map")`a). The preponderance of short branches in southern Honshu, Shikoku, and Kyushu may be due to the radiation of certain species-rich genera in that area such as *Deparia* (`r s_figure("genus-map")`b) and *Dryopteris* (`r s_figure("genus-map")`d).

Furthermore, we identified % apomictic taxa as a strong driver of SES of RPD (Figs. `r figure_num("coeff")`g, `r figure_num("modelfit")`g), and to a slightly lesser extent SES of PD (Figs. `r figure_num("coeff")`f, `r figure_num("modelfit")`f).
Indeed, *Dryopteris* is one such genus with a high rate of apomictic taxa (`r s_figure("apo-genus")`), and other genera with high rates of apomixis (*Cyrtomium*, *Pteris*; `r s_figure("apo-genus")`) also reach greatest richness in the southern main islands (`r s_figure("genus-map")`).
This is consistent with the hypothesis that communities with high proportions of asexual taxa have lower genetic diversity, and is the first time to our knowledge that this has been demonstrated in ferns. 
We verified that this relationship is also observed when analyzed with a phylogeny with branch lengths in units of genetic change (`r ie`, the output of the maximum likelihood phylogenetic analysis, before dating), a more direct measure of genetic diversity (`r s_figure("modelfit-not-ult")`).
While temperature is also correlated with % apomictic taxa and therefore cannot be discounted, conditional AIC indicates a better fit of the reproductive model than the environmental model for SES of RPD (`r s_table("aic")`). 
Ultimately, it is likely that temperature, in combination with other environmental conditions, is driving the predominance of apomictic taxa: @Tanaka2014 also found that the proportion of apomictic taxa reaches a maximum at ca. 34.8 °N latitude and increases with temperature in Japan (their study was limited to the main islands and did not include the smaller southern islands). 
@Tanaka2014 asserted that one limiting factor affecting establishment of apomictic ferns at high latitudes or cold climates is the large cell size of apomicts, which are typically polyploid. 
As fern gametophytes are only a single cell-layer in thickness with no cuticle or stomata, it is expected they would be particularly susceptible to freezing damage induced by larger cells with higher water content; however, no experimental studies have tested these hypotheses to our knowledge.

@@

The variation in distribution of different biodiversity metrics is reflected by the bioregions analysis, which tended to group cells with similar biodiversity values (`r figure("div-by-tax-region")`). 
Interestingly, the bioregions identified here correspond roughly to major vegetation zones previously categorized on an *ad-hoc* basis; bioregion 1 corresponds to the cool or subartic deciduous type, bioregion 2 to the warm evergreen type, and bioregions 3 and 4 to the subtropical rainy type [@Shimizu2014]. 
Furthermore, the border between bioregions 1 and 2 corresponds to the "*Gleichenia japonica* Line" proposed by @Nakaike1983, which he proposed based on the distribution of *Diplopterygium glaucum* (Houtt.) Nakai (= *Gleichenia japonica* Spreng.) as an indicator of evergreen broad-leaf forests in Japan.
This suggests that ferns are useful as bioindicators in Japan. 
Although some studies have emphasized the importance of deep-sea straits (the Tsugaru Strait separating Hokkaido from Honshu and the Tokara Strait separating the southern islands from Kyushu) in structuring biological communities in Japan [@Millien-Parra1999; @Kubota2014], these did not play a major role in structuring fern bioregions. 
The Tsugaru Strait had no relation to any of the bioregions, and although the Tokara Strait somewhat splits phylogenetic bioregions 2 and 3, there is some overlap between them on Yakushima and Amami Isl. (`r figure("bioregions")`b). 
The weak effect of these straits as barriers makes sense given the presumably high dispersal abilities of ferns, which disperse via tiny spores carried on the wind [@Tryon1970].
Rather, fern bioregions in Japan broadly separate along gradients of temperature and precipitation; in particular, precipitation splits taxonomic bioregion 3 (mostly the Ryukyu islands; wetter) from bioregion 4 (mostly the Ogasawara islands; drier) (`r s_figure("bioregions-env")`).

@@

We found that grid-cells with significantly high amounts of biodiversity are better protected than the baseline protection rate at the medium protection level across all biodiversity metrics, and better than the baseline rate at the high protection level for PE (`r figure("conserv-status")`a). 
The high protection status of grid-cells with high PE may be due to an emphasis on conserving endemic plant and animal species in Japanese southern islands such as Iriomote and Amami [@kankyosho2021b], which include many grid-cells with high PE (`r s_figure("protected-map")`d).
One possible reason for the generally high level of protection overall is that in the past, some ferns have been included in lists of rare and endemic plants for consideration in designation of conservation areas [A. Ebihara, pers. comm.; @kankyosho2021b]. 
Also, (as suggested previously) ferns may serve as bioindicators reflecting overall levels of biodiversity, particularly for other plant groups.

@@

Our analysis shows that the threat to fern biodiversity posed by Japanese deer differs strongly by biodiversity type; although grid-cells with significantly high FD, PD, and PE tend to have lower rates of deer occupancy compared to the baseline rate, grid-cells with high richness experience much greater co-occurrence with deer than the baseline rate (`r figure("conserv-status")`b).
This pattern is due to the fact that Japanese deer do not occur in the southern islands, which house high amounts of PD and PE; rather, areas with high species richness are primarily located in southern Honshu, Shikoku, and Kyushu, where deer expansion has been intense (`r s_figure("protected-map")`a).
This result highlights the importance of using multiple metrics of biodiversity for establishing conservation priorities.

@@

The extreme evolutionary distinctiveness of the southern islands is both a major feature of the Japanese fern flora and a challenge for analysis. 
The tropical lineages occurring only in these islands are likely responsible for the comparatively low (`r ie`, clustered) values of PD observed on the main islands (`r figure("rand-diversity")`a).
One way to account for this bimodal distribution pattern would be to use a restricted null model that does not allow all taxa to disperse anywhere within the study area [@Mishler2020]. 
However, ferns are known to have high dispersal abilities [@Tryon1970; @Barrington1993; @Moran2001]. 
In our dataset, the taxon with the greatest latitudinal distribution (`r taxon_with_widest_lat$text`) spans nearly the entire region (`r taxon_with_widest_lat$min_lat`°N to `r taxon_with_widest_lat$max_lat`°N), and `r n_widespread_taxa` taxa have latitudinal distributions spanning the Tokara strait in the south (29°N) to the Tsugaru strait in the north (41.5°N). 
Therefore, it seems reasonable to assume that ferns are capable of dispersing across the entire study area.
Furthermore, it should be noted that the southern islands are all low-altitude (not exceeding 1,000 m).
If high elevation habitats were available, the southern islands would be expected to share more species in common with areas further north.
For example, Taiwan, which is just east of the southernmost Japanese islands and harbors high mountains (> 3,000 m), shares many fern species with the main Japanese islands that are not found in southern Japanese islands.
This supports that the sharp taxonomic turnover between southern and main islands is more likely due to ecological factors than dispersal limitation.

@@

```{r raw-diversity-make}
# Richness
a <- ggplot() +
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = richness), color = "transparent") +
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(A)** Richness</span>", hjust = "inward", vjust = "inward", color = "white") + 
  labs(fill = "No. taxa") +
  scale_fill_scico(palette = "lajolla") + 
  scale_x_continuous(breaks = c(125, 135, 145)) +
  scale_y_continuous(breaks = c(25, 35, 45)) +
  map_theme()

# Raw PD
b <- ggplot() +
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = pd_obs), color = "transparent") +
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(B)** PD</span>", hjust = "inward", vjust = "inward", color = "white") +
  labs(fill = "Relative\nbranch\nlength") +
  scale_fill_scico(palette = "lajolla") + 
  scale_x_continuous(breaks = c(125, 135, 145)) +
  scale_y_continuous(breaks = c(25, 35, 45)) +
  map_theme() +
  theme(axis.text.y = element_blank())

>>>>

We used data compiled by @Ebihara2019b, which classifies each fern taxon as sexual, apomictic, mixed (both sexual and apomictic modes known from the same taxon), or unknown. 
We calculated the percentage of apomictic species in each grid-cell as the sum of apomictic and mixed taxa divided by total number of taxa (`r s_figure("envir")`B).

@@

The cleaned list of specimens included `r n_specimens` specimens representing `r n_specimens_taxa` taxa (`r n_specimens_taxa_percent`). 
The distribution of collection locations was highly skewed, with > 3,000 specimens per grid-cell northwest of Tokyo in Saitama prefecture, and many fewer elsewhere (`r s_figure("abundance")`A). 
Despite this, sampling redundancy was generally good [> 0.3; @Gonzalez-Orozco2014b] throughout the country (mean `r mean_redundancy`, final community data matrix; errors are SD unless mentioned otherwise) (`r s_figure("abundance")`B). 
The final community data matrix included `r ncol(comm_ferns)` taxa and `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` grid-cells. `r n_excluded_cells` grid-cells were excluded due to low redundancy (< 0.1); these were mostly from Hokkaido and coastal areas.

@@

Taxonomic richness is lowest on the northernmost island of Hokkaido, then increases with decreasing latitude until reaching a maximum in southern Kyushu, then decreases again in smaller islands further south (`r figure("raw-diversity")`A, `r s_figure("raw-lat")`A). 
Observed values of PD and FD were highly correlated with richness as expected (`r figure("raw-diversity")`B, C), showing an initial steep slope that gradually starts to level off at higher richness (`r s_figure("raw-biplot")`A, C). 
Observed values of RPD and RFD were not linearly related to richness; however, they showed high variance at low species richness, and low variance at high species richness (`r s_figure("raw-biplot")`B, D).

@@

Grid-cells with significantly low PD were almost all located on the main islands of Japan (Honshu, Kyushu, and Shikoku; `r figure("rand-diversity")`A). 
A smaller number of cells had significantly high PD, and these were located mostly on small southern (Ryukyu, Izu, Ogasawara) islands or coastal areas (Kii Peninsula). 
Grid-cells with significantly low RPD were mostly observed in the southern main islands, particularly in and around Kyushu (`r figure("rand-diversity")`B). 
Grid-cells with significantly high RPD are located both in the small southern islands and northern Honshu and Hokkaido. 
Results of the randomization test for functional diversity were broadly similar to phylogenetic diversity (`r figure("rand-diversity")`C, D).

@@

All models were significant relative to null models without any fixed effects (`r s_table("lrt-null")`). 
The effect of temperature dominated (had the greatest absolute *t*-value) in environmental models (`r figure("coeff")`A--E), and % apomictic taxa dominated in reproductive models (`r figure("coeff")`F--G); temperature and % apomictic taxa were significant in each model where they were included, whereas other predictors (area, precipitation, precipitation seasonality) were not significant in some models. 
Richness showed a hump-shaped relationship with temperature, reaching maximum richness at intermediate temperatures (`r figure("modelfit")`A). 
Phylogenetic and functional diversity showed either moderately (SES of PD, SES of FD) to strongly (SES of RPD, SES of RFD) hump-shaped relationships with temperature (`r figure("modelfit")`B--E). 
SES of RPD showed a clear, linear decline with % apomictic taxa (`r figure("modelfit")`G); a similar pattern was detected in SES of PD, but with wider confidence intervals (`r figure("modelfit")`F). 
Conditional AIC was lower (indicating better model fit) in the reproductive model compared to the environmental model for SES of RPD and SES of PD (`r s_table("aic")`).

@@

Clustering analysis grouped grid-cells into `r english2(k_taxonomy$optimal$k)` or `r english2(k_phylogeny$optimal$k)` bioregions (`r s_figure("k-plot")`) on the basis of taxonomic or phylogenetic distances, respectively. 
The vast majority of grid-cells fell to a subset of these bioregions: using taxonomic distances, they include bioregion 1 (Hokkaido and northern Honshu), bioregion 2 (southern Honshu, Kyushu, and Shikoku), bioregion 3 (Ryukyu Islands), and bioregion 4 (Ogasawara Islands) (`r figure("bioregions")`A; `r s_figure("dendrogram")`A). 
Phylogenetic distances produced similar results, but the Ryukyu and Ogasawara Islands merged into a single bioregion (3), and bioregion 1 extended slightly further south (`r figure("bioregions")`B; `r s_figure("dendrogram")`B). 
The other much smaller bioregions ("singletons" including only a single grid-cell each) are likely artifacts of insufficiently low taxon richness for robust clustering, and are not discussed further.

@@

The percentage of protected area in grid-cells with significantly high biodiversity range from `r min_protected_total` to `r max_protected_total` total, including `r min_protected_high` to `r max_protected_high` with high status and `r min_protected_med` to `r max_protected_med` with medium status (`r figure("conserv-status")`A). 
Cells with significantly high biodiversity had a greater percentage of area protected than the baseline rate for Japan at the medium protection status (`r scales::percent(total_percent_protected_medium, accuracy = 0.1)`) across all biodiversity metrics.
In contrast, cells with significantly high biodiversity had a lower percentage of area protected than the baseline rate for Japan at the high protection status (`r scales::percent(total_percent_protected_high, accuracy = 0.1)`) for most biodiversity metrics.
Only cells with significantly high PE had a higher percentage area protected (`r best_protected_high` high, `r best_protected_med` medium, `r best_protected_total` total) than the baseline rate at both levels of protection.

The range of Japanese deer in Japan increased from `r deer_overall[["1978"]]$area` (`r deer_overall[["1978"]]$percent`) in 1978 to `r deer_overall[["2003"]]$area` (`r deer_overall[["2003"]]$percent`) in 2003, and may be as high as `r deer_overall$estimated$area` (`r deer_overall$estimated$percent`) according to model estimates.
The percentage of area occupied by deer in grid-cells with significantly high biodiversity increased from `r deer_in_high_biodiv[["1978"]]$percent` in 1978 to `r deer_in_high_biodiv[["2003"]]$percent` in 2003 when averaged across biodiversity types.
Cells with high FD, PD, or PE tended to have less deer occupation compared to the baseline rate (`r figure("conserv-status")`B).
However, cells with high species richness consistently exceeded the baseline rate of deer presence, with up to `r deer_highest_danger$estimated$percent` occupied by deer in the estimated range dataset.

@@

The latitudinal hump-shaped pattern of species richness observed in ferns here (Figs. `r figure_num("raw-diversity")`a, `r s_figure_num("raw-lat")`a) is well-documented [@Kusumoto2017; @Ebihara2019b], and similar patterns have been observed in deciduous broadleaved trees and conifers [@Shiono2015], and niche-modeled distributions of woody plants in Japan [@Fukaya2020]. 
Our spatial modeling analysis indicates that this is strongly linked to temperature, with maximum richness occurring at intermediate temperatures (`r figure("modelfit")`A).
Previous studies have identified temperature, water availability [@Bickford2006; @Qian2012], or a combination of these [@Nagalingum2015; @Link-Perez2018; @Qian2021] to be important drivers of spatial fern diversity.
A hump-shaped pattern of richness along elevation (maximum richness at intermediate elevations) has been frequently observed in ferns on mountains in tropical or sub-tropical areas [@Kessler2011a; @Suissa2021]. 
Hypotheses to explain such a pattern include purely geometrical constraints as well as biological hypothesis positing maximally favorable environmental conditions at mid-elevations [@Colwell2004a]. 
Another reason for the greater richness on the main islands compared to southern islands in Japan is that the main islands harbor higher elevational variation that may support a wider range of taxa than is present in the southern islands.

The overwhelming pattern characterizing phylogenetic and functional diversity (PD and FD) is the high diversity of southern islands (Ryukyu and Ogasawara Isl.) (`r figure("rand-diversity")`A, C). 
This is almost certainly due to the presence of primarily tropical lineages that do not occur in other parts of the country; `r n_southern_genera` genera and `r n_southern_families` families only occur south of 30.1° latitude (just south of Yakushima Isl.). 
The similarity in patterns of PD and FD is likely due to the moderate degree of phylogenetic signal present in the traits used in our analysis (Tables `r s_table_num("phy-sig")`, `r s_table_num("phy-sig-binary")`), and suggests that in ferns, phylogeny can be used as a reasonable stand-in for functional diversity. 
Given their high phylogenetic diversity, it is perhaps not surprising that southern islands also host high amounts of phylogenetic endemism; indeed, the vast majority of cells with significant levels of PE, as well as paleoendemic cells, are located in the southern islands (Figs. `r figure_num("endemism")`, `r s_figure_num("endemism-restricted")`). 
This is clearly due to the small land area of these islands, which are the only areas with a subtropical climate in Japan.

Another striking pattern revealed here is the predominance of long branches in the northern half of Honshu and the southern islands, and the predominance of short branches in southern Honshu, Shikoku, and Kyushu (`r figure("rand-diversity")`B). 
The long branches in the southern islands are likely being driven by the tropical lineages mentioned previously; those in the north may be due to refugial lineages that are distantly related to others and become more species rich at high-latitudes, such as *Equisetum* (`r s_figure("genus-map")`A). The preponderance of short branches in southern Honshu, Shikoku, and Kyushu may be due to the radiation of certain species-rich genera in that area such as *Deparia* (`r s_figure("genus-map")`B) and *Dryopteris* (`r s_figure("genus-map")`D).

Furthermore, we identified % apomictic taxa as a strong driver of SES of RPD (Figs. `r figure_num("coeff")`G, `r figure_num("modelfit")`G), and to a slightly lesser extent SES of PD (Figs. `r figure_num("coeff")`F, `r figure_num("modelfit")`F).
Indeed, *Dryopteris* is one such genus with a high rate of apomictic taxa (`r s_figure("apo-genus")`), and other genera with high rates of apomixis (*Cyrtomium*, *Pteris*; `r s_figure("apo-genus")`) also reach greatest richness in the southern main islands (`r s_figure("genus-map")`).
This is consistent with the hypothesis that communities with high proportions of asexual taxa have lower genetic diversity, and is the first time to our knowledge that this has been demonstrated in ferns. 
We verified that this relationship is also observed when analyzed with a phylogeny with branch lengths in units of genetic change (`r ie`, the output of the maximum likelihood phylogenetic analysis, before dating), a more direct measure of genetic diversity (`r s_figure("modelfit-not-ult")`).
While temperature is also correlated with % apomictic taxa and therefore cannot be discounted, conditional AIC indicates a better fit of the reproductive model than the environmental model for SES of RPD (`r s_table("aic")`). 
Ultimately, it is likely that temperature, in combination with other environmental conditions, is driving the predominance of apomictic taxa: @Tanaka2014 also found that the proportion of apomictic taxa reaches a maximum at ca. 34.8 °N latitude and increases with temperature in Japan (their study was limited to the main islands and did not include the smaller southern islands). 
@Tanaka2014 asserted that one limiting factor affecting establishment of apomictic ferns at high latitudes or cold climates is the large cell size of apomicts, which are typically polyploid. 
As fern gametophytes are only a single cell-layer in thickness with no cuticle or stomata, it is expected they would be particularly susceptible to freezing damage induced by larger cells with higher water content; however, no experimental studies have tested these hypotheses to our knowledge.

@@

The variation in distribution of different biodiversity metrics is reflected by the bioregions analysis, which tended to group cells with similar biodiversity values (`r figure("div-by-tax-region")`). 
Interestingly, the bioregions identified here correspond roughly to major vegetation zones previously categorized on an *ad-hoc* basis; bioregion 1 corresponds to the cool or subartic deciduous type, bioregion 2 to the warm evergreen type, and bioregions 3 and 4 to the subtropical rainy type [@Shimizu2014]. 
Furthermore, the border between bioregions 1 and 2 corresponds to the "*Gleichenia japonica* Line" proposed by @Nakaike1983, which he proposed based on the distribution of *Diplopterygium glaucum* (Houtt.) Nakai (= *Gleichenia japonica* Spreng.) as an indicator of evergreen broad-leaf forests in Japan.
This suggests that ferns are useful as bioindicators in Japan. 
Although some studies have emphasized the importance of deep-sea straits (the Tsugaru Strait separating Hokkaido from Honshu and the Tokara Strait separating the southern islands from Kyushu) in structuring biological communities in Japan [@Millien-Parra1999; @Kubota2014], these did not play a major role in structuring fern bioregions. 
The Tsugaru Strait had no relation to any of the bioregions, and although the Tokara Strait somewhat splits phylogenetic bioregions 2 and 3, there is some overlap between them on Yakushima and Amami Isl. (`r figure("bioregions")`B). 
The weak effect of these straits as barriers makes sense given the presumably high dispersal abilities of ferns, which disperse via tiny spores carried on the wind [@Tryon1970].
Rather, fern bioregions in Japan broadly separate along gradients of temperature and precipitation; in particular, precipitation splits taxonomic bioregion 3 (mostly the Ryukyu islands; wetter) from bioregion 4 (mostly the Ogasawara islands; drier) (`r s_figure("bioregions-env")`).

@@

We found that grid-cells with significantly high amounts of biodiversity are better protected than the baseline protection rate at the medium protection level across all biodiversity metrics, and better than the baseline rate at the high protection level for PE (`r figure("conserv-status")`A). 
The high protection status of grid-cells with high PE may be due to an emphasis on conserving endemic plant and animal species in Japanese southern islands such as Iriomote and Amami [@kankyosho2021b], which include many grid-cells with high PE (`r s_figure("protected-map")`D).
One possible reason for the generally high level of protection overall is that in the past, some ferns have been included in lists of rare and endemic plants for consideration in designation of conservation areas [A. Ebihara, pers. comm.; @kankyosho2021b]. 
Also, (as suggested previously) ferns may serve as bioindicators reflecting overall levels of biodiversity, particularly for other plant groups.

@@

Our analysis shows that the threat to fern biodiversity posed by Japanese deer differs strongly by biodiversity type; although grid-cells with significantly high FD, PD, and PE tend to have lower rates of deer occupancy compared to the baseline rate, grid-cells with high richness experience much greater co-occurrence with deer than the baseline rate (`r figure("conserv-status")`B).
This pattern is due to the fact that Japanese deer do not occur in the southern islands, which house high amounts of PD and PE; rather, areas with high species richness are primarily located in southern Honshu, Shikoku, and Kyushu, where deer expansion has been intense (`r s_figure("protected-map")`A).
This result highlights the importance of using multiple metrics of biodiversity for establishing conservation priorities.

@@

The extreme evolutionary distinctiveness of the southern islands is both a major feature of the Japanese fern flora and a challenge for analysis. 
The tropical lineages occurring only in these islands are likely responsible for the comparatively low (`r ie`, clustered) values of PD observed on the main islands (`r figure("rand-diversity")`A).
One way to account for this bimodal distribution pattern would be to use a restricted null model that does not allow all taxa to disperse anywhere within the study area [@Mishler2020]. 
However, ferns are known to have high dispersal abilities [@Tryon1970; @Barrington1993; @Moran2001]. 
In our dataset, the taxon with the greatest latitudinal distribution (`r taxon_with_widest_lat$text`) spans nearly the entire region (`r taxon_with_widest_lat$min_lat`°N to `r taxon_with_widest_lat$max_lat`°N), and `r n_widespread_taxa` taxa have latitudinal distributions spanning the Tokara strait in the south (29°N) to the Tsugaru strait in the north (41.5°N). 
Therefore, it seems reasonable to assume that ferns are capable of dispersing across the entire study area.
Furthermore, it should be noted that the southern islands are all low-altitude (not exceeding 1,000 m).
If high elevation habitats were available, the southern islands would be expected to share more species in common with areas further north.
For example, Taiwan, which is just east of the southernmost Japanese islands and harbors high mountains (> 3,000 m), shares many fern species with the main Japanese islands that are not found in southern Japanese islands.
This supports that the sharp taxonomic turnover between southern and main islands is more likely due to ecological factors than dispersal limitation.

@@

```{r raw-diversity-make}
# Richness
a <- ggplot() +
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = richness), color = "transparent") +
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**A** Richness</span>", hjust = "inward", vjust = "inward", color = "white") + 
  labs(fill = "No. taxa") +
  scale_fill_scico(palette = "lajolla") + 
  scale_x_continuous(breaks = c(125, 135, 145)) +
  scale_y_continuous(breaks = c(25, 35, 45)) +
  map_theme()

# Raw PD
b <- ggplot() +
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = pd_obs), color = "transparent") +
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**B** PD</span>", hjust = "inward", vjust = "inward", color = "white") +
  labs(fill = "Relative\nbranch\nlength") +
  scale_fill_scico(palette = "lajolla") + 
  scale_x_continuous(breaks = c(125, 135, 145)) +
  scale_y_continuous(breaks = c(25, 35, 45)) +
  map_theme() +
  theme(axis.text.y = element_blank())

# Raw FD
c <- ggplot() + 
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = fd_obs), color = "transparent") +
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**C** FD</span>", hjust = "inward", vjust = "inward", color = "white") + 
  labs(fill = "Relative\nbranch\nlength") +
  scale_fill_scico(palette = "lajolla") + 
  scale_x_continuous(breaks = c(125, 135, 145)) +
  scale_y_continuous(breaks = c(25, 35, 45)) +
  annotation_scale(location = "tl", width_hint = 0.2, pad_y = unit(0.8, "cm")) +
  map_theme() +
  theme(axis.text.y = element_blank())
